# Functional programming - Exercises

## Anonymous functions

Q1. Given a function, like "mean", match.fun() lets you find a function. Given a
function, can you find its name? Why doesn’t that make sense in R?

SKIP

Q2. Use lapply() and an anonymous function to find the coefficient of variation (the
standard deviation divided by the mean) for all columns in the mtcars dataset.

```{r coeff_var}
lapply(mtcars, function(x) sd(x) / mean(x))
```

Q3. Use integrate() and an anonymous function to find the area under the curve for
the following functions. Use Wolfram Alpha to check your answers. 
y = x ^ 2 - x, x in [0, 10] 
y = sin(x) + cos(x), x in [-π, π] 
y = exp(x) / x, x in [10, 20]

```{r}
integrate(function(x) x ^ 2 - x, 0, 10)
integrate(function(x) sin(x) + cos(x), -pi, pi)
integrate(function(x) exp(x) / x, 10, 20)
```

Q4. A good rule of thumb is that an anonymous function should fit on one line and
shouldn’t need to use {}. Review your code. Where could you have used an
anonymous function instead of a named function? Where should you have used a
named function instead of an anonymous function?

SKIP

## Closures

Q1. Why are functions created by other functions called closures?

Because they enclose the environment of the parent function, and can thus access
variables in that environment.

Q2. What does the following statistical function do? What would be a better name
for it? (The existing name is a bit of a hint.)

It returns a function that performs a Box-Cox transformation of the input.

```r
bc <- function(lambda) { 
  if (lambda == 0) { 
    function(x) log(x) 
  } else { 
    function(x) (x ^ lambda - 1) / lambda }
}
```

Q3. hat does `approxfun()` do? What does it return?

It returns a function that does linear interpolation of the data given as input:

```{r approxfun}
approxfun(x = c(1, 2), y =  c(1, 2))
```

Q4. What does `ecdf()` do? What does it return?

It calculates the empirical cumulative density function of a given input and
returns a function. The returned function takes a value as input, and returns
the estimated cumulative density of the original data at that value.

```{r ecdf}
ecdf(1:10)(5)
```

Q5. Create a function that creates functions that compute the ith central moment
of a numeric vector. You can test it by running the following code:

```r
m1 <- moment(1)
m2 <- moment(2)

x <- runif(100) 
stopifnot(all.equal(m1(x), 0)) 
stopifnot(all.equal(m2(x), var(x) * 99 / 100))
```

```{r central_moment}
moment <- function(n) {
  function(x) {
    mean((x - mean(x)) ^ n)
  }
}
```

6. Create a function `pick()` that takes an index, `i`, as an argument and returns
a function with an argument `x` that subsets `x` with `i`.

```r
lapply(mtcars, pick(5)) # should do the same as this 
lapply(mtcars, function(x) x[[5]])
```

```{r pick}
pick <- function(i) {
  function(x) {
    x[[i]]
  }
}

identical(
  lapply(mtcars, pick(5)),
  lapply(mtcars, function(x) x[[5]])
)
```

## List of functions

Q1. Implement a summary function that works `like base::summary()`, but uses a list of
functions. Modify the function so it returns a closure, making it possible to
use it as a function factory.

Q2. Which of the following commands is equivalent to `with(x, f(z))`? 
`x$f(x$z)`
`f(x$z)`
`x$f(z)`
`f(z)`
It depends.


## Case study: numerical integration

Q1. Instead of creating individual functions (e.g., `midpoint()`, `trapezoid()`,
`simpson()`, etc.), we could store them in a list. If we did that, how would that
change the code? Can you create the list of functions from a list of
coefficients for the Newton-Cotes formulae?

Q2. The trade-off between integration rules is that more complex rules are slower to
compute, but need fewer pieces. For `sin()` in the range `[0, π]`, determine the
number of pieces needed so that each rule will be equally accurate. Illustrate
your results with a graph. How do they change for different functions? `sin(1 /
x^2)` is particularly challenging.
